Texture mapping 3D planar models of indoor environments with noisy camera poses

### 2013-01-10 spie/paper.tex
new paper draft

  Automated 3D modeling of building interiors is used in applications
  such as virtual reality and environment mapping. Texturing these
  models allows for photorealistic visualizations of the data
  collected by such modeling systems. Camera poses obtained by these
  systems often suffer from inaccuracies, resulting in visible
  discontinuities when successive images are projected adjacently onto
  a surface for texturing. Existing methods to stitch images together
  are often computationally expensive and work independently of pose
  estimates and geometry information. We propose a method to refine
  camera poses using both existing estimates and geometry information,
  followed by two different methods to composite images together,
  based on the uniformity of available images. The effectiveness of
  our methods is demonstrated on a number of different indoor
  environments.
\end{abstract}

@@

In this paper, we work with data obtained from a backpack-mounted
system, carried by an ambulatory human. Such a system provides
advantages over more common wheel-mounted systems in terms of agility
and portability, but results in much higher localization error
\cite{liu2010indoor}. As a result, common methods for texture mapping
generally produce poor results, leading to the development of the
approaches contained in this paper. Besides localization information
and images, environment geometry is required as well for texture
mapping. In this paper we work mainly with models obtained by fitting
planar surfaces to point clouds generated by the backpack system
\cite{sanchez2012point}. We have also developed an appropriate method
to work with near-planar triangular meshes, as described in Appendix
A.

Though our texture mapping procedure was designed with the
aforementioned system in mind, it is applicable to other systems,
provided the required inputs of planar geometry, images with rough
extrinsic pose estimates, and either known or estimated intrinsic
camera calibration matrices. An important benefit of our method is its
relatively low complexity and runtime, and independence of many
steps, allowing for quick tuning of parameters and easy
accomodation of new input data given updates in localization or
model-generation algorithms.

The remainder of the paper is organized as follows. Section
\ref{sec:simpleTextureMapping} explains how input images are projected
onto our geometry and simple approaches towards texturing. Section
\ref{sec:existingApproaches} covers more sophisticated existing
approaches to image matching and stitching, and their performance on
our datasets. Section \ref{sec:2dAlignment} contains our approach
towards image alignment, followed by Section
\ref{sec:imageCompositing}, which describes compositing
approaches. Section \ref{sec:resultsAndConclusions} contains further
results and conclusions, as well as areas of ongoing and future work.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The geometry of the texture mapping process for a plane is shown in
Figure \ref{fig:projection}.  As described earlier, we are provided
with a set of $M$ images to texture the target plane. Each image has a
camera matrix $P_i$ for $i=1..M$, which translates a 3D point in the
world coordinate system to a 2D point or pixel in image $i$'s
coordinates. A camera matrix $P_i$ is composed of the camera's
intrinsic parameters, such as focal length and image center, as well
as extrinsic parameters which specify the rotation and translation of
the camera's position in 3D world coordinates at the time that image
$i$ is taken. These extrinsic parameters are determined by the
backpack hardware and localization algorithms \cite{chen2010indoor,
  liu2010indoor, kua2012loopclosure} and are quite noisy.

Because our backpack system takes photos at a rate of 5 Hz, thousands
of images are available for texturing each surface in our model. Our
goal in designing a texture mapping process is to decide which of
these images should be used, and where their contents should map onto
the texture, in order to minimize any visual discontinuities or seams
that would suggest that the plane's final texture is not composed of a
single continuous image.

@@

\subsection{Tile-based Texture Mapping}
\label{sec:tileBasedMapping}
Ignoring the fact that the camera matrices $P_{1..M}$ are inaccurate,
we can texture the plane by discretizing it into small square tiles,
in our case 5 pixels across, and choosing an image to texture each
tile.

@@

Once we have a list of candidate images for $t$, we define a scoring
function in order to objectively select the best image for texturing
$t$. Since camera pose errors compound over distance, we wish to
minimize the distance between cameras and the surfaces they
texture. Additionally, we desire images that are projected
perpendicularly onto the plane, maximizing the resolution and amount
of useful texture available in their projections, as well as
minimizing any parallax effects due to real-world geometry not
accurately represented by our digital model. In other words, we wish
to minimize the angle between the tile's normal vector and the camera
axis for images selected for texture mapping. These two criteria can
be met by maximizing the function $\frac{1}{d} (-1 \cdot \vec{c})
\cdot \vec{n}$ as shown in Figure
\ref{fig:scoringFunction}. Specifically, $d$ is the distance between
the centers of a camera and a tile, and $\vec{n}$ and $\vec{c}$ are
the directions of the plane's normal and the camera axis respectively.

\begin{figure}
  \centering 
  \subfloat[][]{\includegraphics[width=3in, height=0.97in]{wall2_naive.pdf}}
  \subfloat[][]{\includegraphics[width=3in, height=0.97in]{wall2_naive_shift.pdf}}

  \centering
  \subfloat[][]{\includegraphics[width=3in, height=0.97in]{wall2_cache_shift.pdf}}
  \subfloat[][]{\includegraphics[width=3in, height=0.97in]{wall2_shortest_shift.pdf}}

  \centering 
  \subfloat[][]{\includegraphics[width=3in, height=0.97in]{wall2_cache_shift_blend.pdf}}
  \subfloat[][]{\includegraphics[width=3in, height=0.97in]{wall2_shortest_shift_blend.pdf}}
  \caption{(a) Tile-based texturing. (b) Tile-based texturing after
    image alignment. (c) Tile-based texturing after image alignment
    with caching. (d) Shortest path texturing after image
    alignment). (e) Same as (c) with blending. (f) Same as (d) with
    blending.}
  \label{fig:compareAll}
\end{figure}

As Figure \ref{fig:compareAll}(a) demonstrates, this approach leads to
the best texture for each tile independently, but results in many
image boundaries with abrupt discontinuities between tiles, due to
significant misalignment between images. Given our high number of
input images, it makese sense to further refine the image selection
procedure, as performed in Section \ref{sec:imageCompositing}, but for
our solution to be more robust, we first perform image alignment in
order maximize the accuracy of each image.

@@

Feature detection and matching works best when multiple unique visual
references exist in the environment and are present within multiple
images. In contrast, our indoor environments have a high prevalence of
bare surfaces, as well as repeating textures that cause difficulty in
disambiguating features. This lack of strong reference points results
in high uncertainty when matching images together.

Additionally, our datasets often contain long chains of images, which
leads to error accumulation when image correspondences are not
accurate. For example, when matching a long chain of images through
homographies, a pixel in the $nth$ image must be translated into the
first image's coordinates by multiplying by the $3\times3$ matrix $H_1
H_2 H_3 ... H_n$. Any error in one of these homography matrices is
propagated to all further images, resulting in drift. Figure
\ref{fig:mosaic3D}(b) shows the output of the AutoStitch software
package, which performs homography-based image mosaicing
\cite{autostitch}. Even with many features spread across this surface,
the mosaicing produces errors that cause straight lines to appear
distorted, despite the fact that it was generated after careful hand
tuning. Many areas with fewer visual features simply failed outright
using this approach.

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=3in]{Graph_crop.pdf}}
  \subfloat[][]{\includegraphics[width=3in]{panoMy.jpg}}

  \caption{Texture alignment via (a) the graph-based localization
    refinement algorithm from \cite{chen2010indoor} and (b) image
    mosaicing.}
  \label{fig:mosaic3D}
\end{figure}

Image projections can also be aligned by iteratively adjusting camera
poses to maximize matches. This is generally done over 6 degrees of
freedom \cite{liu2010indoor}. The approach in the cited paper works
well locally, given high feature density, but over larger areas
results again in error accumulation and drift, especially in
feature-sparse areas, as shown in Figure
\ref{fig:mosaic3D}(a). Furthermore, it adjusts image locations using
an iterative process, which results in high runtime.

@@

All of our following calculations and alignments are performed in 2D,
partly for efficiency, and partly because the nature of our input data
is such that localization error chiefly occurs in two dimensions, in
the same plane as the surface being projected onto. Recall that our
data acquisition system is backpack-mounted, with cameras facing to
the sides of the operator. The operator makes efforts to walk parallel
to walls, and the localization and model-generation algorithms that
provide our input often conform the operator's path to be straight and
parallel to detected surfaces \cite{kua2012loopclosure,
  sanchez2012point}. Since the operator walks in as upright a position
as possible, errors in roll, yaw, and the operator's distance from
parallel surfaces are minimal. Thus, our highest errors stem from
uncertainty in the operator's pitch, equivalent to rotation around the
camera axes, as well as location along surfaces. These equate to 2D
rotation and translation in a surface's plane.

\subsection{Geometry-based Alignment}
\label{sec:geometryAlignment}
After computing each image's projection onto the target surface, as
described in Section \ref{sec:simpleTextureMapping}, we detect lines
in the projections using Hough transforms. Experience and intuition
show that walls in indoor environments contain linear features that
are either horizontal or vertical, often corresponding to doors,
windows, posters, etc. Thus, when texturing walls, we rotate images in
2D such that dominant lines are made to be horizontal or
vertical. This can be further extrapolated by orienting lines with a
wall's boundaries, for example in areas with a slanted roof.

\begin{figure}
  \centering
  \includegraphics[height=1.5in]{geometryAlign_planes.png}
  \includegraphics[height=1.5in]{geometryAlignment.jpg}
  \caption{(a) When texturing the red surface, it makes sense to align images to surface boundaries and intersections. (b) Red lines are geometry-based lines, while green lines are detected in the image via Hough transform. Above is the original projection, and below is the image adjusted for alignment.}
  \label{fig:geometryAlignment}
\end{figure}

At this point in time, image occlusions have not been accounted
for. As a result, some image projections contain texture that should
project to an adjacent surface, generally with a linear boundary where
the two surfaces meet. If this linear boundary is detected by Hough
transform, the image is rotated and shifted in 2D such that the visual
boundary between two surfaces matches the physical boundary in our
digital model. An example of such an adjustment is in Figure
\ref{fig:geometryAlignment}

To perform the rotation and alignment, one method would be to detect
all line matches and solve for an optimal homography. In nearly all
cases however, we find that a single 2D rotation is sufficient to
properly align lines in an image. We also rarely have more than one
linear geometry feature to match to in each orthogonal
direction. Thus, for a rotation angle, we simply take all angles
between image-based and geometry-based lines under 5 degrees, and
average them. After the image is rotated, we find the horizontal and
vertical image-based and geometry-based line segments with minimum
orthogonal distance between them. For both distances, if it is under 5
mm, we apply the corresponding shift, and indicate that the image was
aligned to geometry in one or both dimensions, which reduces further
shifting in Section \ref{sec:robustSIFTFeatureMatching}.

@@

\subsection{2D Feature Alignment}
\label{sec:robustSIFTFeatureMatching}
Our next step is to align overlapping images by searching for
corresponding points between all pairs of overlapping images. We use
SIFT features for their high detection rate, and choose to use feature
alignment rather than pixel or intensity-based alignment due to the
differences in lighting as well as possible occlusion among our
images, both of which feature alignment is less sensitive to
\cite{lowe1999object, mikolajczyk2005performance, szeliski2006image}.
We use SiftGPU \cite{siftgpu} for its high performance on both feature
detection as well as pairwise matching. These matches determine $d^x$
and $d^y$ distances between each pair of features for two image
projections, though these distances may not always be the same for
different features. Since indoor environments often contain repetitive
features such as floor tiles or doors, we need to ensure that
SIFT-based distances are reliable. First, we only perform alignment on
pairs of images that overlap given the original noisy poses. Second,
we discard feature matches that correspond to an image distance
greater than 40 pixels from what the noisy poses estimate. In order to
utilize the remaining feature matches robustly, the RANSAC framework
\cite{fischler1981random} is used to estimate the optimal $d^x_{i,j}$
and $d^y_{i,j}$ distances between two images $i$ and $j$. The RANSAC
framework requires a fitting function and a distance function. For
this application, the fitting function finds the average distance
between matches in a pair of images, and the distance function for a
pair of points is chosen to be the difference between those points'
SIFT match distance and the average distance computed by the fitting
function. We use a 10 pixel threshold, so that SIFT matches are
labeled as outliers if their horizontal or vertical distances are not
within 10 pixels of the average distance computed by the fitting
function.

We now use the calculated $d^x_{i,j}$ and $d^y_{i,j}$ distances
between each pair of images to refine their positions using weighted
linear least squares. There are a total of $M^{2}$ possible pairs of
images, though we only generate distances between images with SIFT
matches. Given these distances and the original image location
estimates, we can solve a least squares problem
($\textrm{min}_{\vec{\beta}} ||A \vec{\beta} - \vec{\gamma}||_2^2 $)
to estimate the location of the images on the plane. The
$M$-dimensional vector $\vec{\beta}$ represents the unknown $x$
location of each image on the plane for $1 \dots M$. The optimal $x$
and $y$ locations are obtained in the same way, so we only consider
the $x$ locations here:

@@

After completing this same process for the $y$ dimension as well, and
making the resultant shifts, our images overlap and match each other
with far greater accuracy. Applying the simple mapping scheme in
Section \ref{sec:tileBasedMapping} to the same wall used in that section results in Figure \ref{fig:compareAll}(b), which has far
fewer seams, though some are still visible.

@@

The result of this caching approach is shown in Figure
\ref{fig:compareAll}(c), where seams are now greatly reduced as
compared to Figure \ref{fig:compareAll}(b).

@@

Figure \ref{fig:dagCreation} demonstrates the construction of a DAG
from overlapping images of a hallway wall. Images are sorted by
horizontal location left to right, and become nodes in a
graph. Directed edges are placed in the graph from left to right
between overlapping images. The weights of these edges are determined
by the SSD cost function. Next, we add two artificial nodes, one start
node representing the left border of the plane, and one end node
representing the right border of the plane. The left(right) artificial
node has directed edges with equal and arbitrary cost $C_0$ to(from)
all images that meet the left(right) border of the plane.

We now solve the shortest path problem from the start node to the end
node. This results in a set of images completely covering the plane
horizontally, while minimizing the cost of seams between images.

@@

After performing linear alpha blending across overlapping regions, our
texture mapping process is complete. Figures \ref{fig:compareAll}(e)
and \ref{fig:compareAll}(f) show the blended versions of Figures
\ref{fig:compareAll}(c) and \ref{fig:compareAll}(d) respectively. It
is clear that the shortest path approach exhibits better alignment and
fewer seams than the tile-caching method, as Figure
\ref{fig:compareAll}(f) has the best visual quality among the textures
in Figure \ref{fig:compareAll}.

\section{Results and Conclusions}
\label{sec:resultsAndConclusions}
Examples of ceilings and floors textured with the tile caching
approach, and walls textured with the shortest path approach, are
displayed in Figure \ref{fig:results}. High resolution texture
comparisons, as well as a walkthrough of a fully textured 3D model are
available in the accompanying video to this paper.

As mentioned earlier, our approach is fairly efficient. The top wall
in Figure \ref{fig:results}(a) is 7543 $\times$ 776 pixels, and spans a
40-meter long wall. Given 41000 input images, it takes approximately a
minute to pick 36 candidate images, followed, by another minute to
perform the shortest path texturing, both on a consumer-grade 2.8GHz
dual-core laptop with 4GB RAM. While not fast enough for real-time
visualization, the procedure is quick enough for quick evaluation
given changes in various internal parameters or modifications to input
data.

\subsection{Future stuff}
For the sake of simplicity, - cases where multiple whatever are few,
but for true generality, should go to 3d adjustments

\begin{figure*}
  \centering
  \subfloat[][]{\includegraphics[width=3in]{finalfloors.jpg}} ~~~~~~~~
  \centering
  \subfloat[][]{\includegraphics[width=3in]{finalceilings.jpg}}

  \centering \subfloat[][]{\includegraphics[height=2in,
    width=3in]{floorcropped.jpg}} ~~~~~~~~ \centering
  \subfloat[][]{\includegraphics[width=3in]{fullmodel.png}}
  \caption{Examples of our final texture mapping output for (a) walls,
    (b) ceilings, (c) an entire floor, (d) a full model.}
  \label{fig:results}
\end{figure*}

\appendix
\section{Working with Near-Planar Geometry}
do stuff and stuff

>>>>

  Automated 3D modeling of building interiors is used in applications
  such as virtual reality and environment mapping. Texturing these
  models allows for photorealistic visualizations of the data
  collected by such modeling systems. Camera poses obtained by these
  systems and used for texturing often suffer from inaccuracies
  however, resulting in visible discontinuities when successive images
  are projected adjacently onto a surface for texturing. Existing
  methods to stitch images together are often computationally
  expensive and work independently of pose estimates and geometry
  information. We propose a method to refine camera poses using both
  existing estimates and geometry information, followed by two
  different methods to composite images together, based on the
  uniformity of available images. The effectiveness of our methods is
  demonstrated on a number of different indoor environments.
\end{abstract}

@@

In this paper, we work with data obtained from a backpack-mounted
system, carried by an ambulatory human. Such a system provides
advantages over more common wheel-mounted systems in terms of agility
and portability, but results in much higher localization error
\cite{liu2010indoor}. As a result, common methods for texture mapping
generally produce poor results, leading to the development of the
approaches contained in this paper. Besides localization information
and images, environment geometry is required as well for texture
mapping. In this paper we work with models obtained by fitting planar
surfaces to point clouds generated by the backpack system
\cite{sanchez2012point}.A.

Though our texture mapping procedure was designed with the
aforementioned system in mind, it is applicable to other systems,
provided the required inputs of planar geometry, images with rough
extrinsic pose estimates, and either known or estimated intrinsic
camera calibration matrices. An important benefit of our method is its
relatively low complexity and runtime, and independence of many steps,
allowing for live tuning of parameters and easy accomodation of new
input data given updates in localization or model-generation
algorithms.

The remainder of the paper is organized as follows. Section
\ref{sec:simpleTextureMapping} explains how input images are projected
onto our geometry and a simple approach towards texturing. Section
\ref{sec:existingApproaches} covers more sophisticated existing
approaches, using image matching and stitching, and their performance
on our datasets. Section \ref{sec:2dAlignment} contains our approach
towards image alignment, followed by Section
\ref{sec:imageCompositing}, which describes our two methods of compositing images. Section \ref{sec:resultsAndConclusions} contains results and conclusions.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The geometry of the texture mapping process for a planar surface is
shown in Figure \ref{fig:projection}.  As described earlier, we are
provided with a set of $M$ images to texture a target plane. Each
image has a camera matrix $P_i$ for $i=1..M$, which translates a 3D
point in the world coordinate system to a 2D point or pixel in image
$i$'s coordinates. A camera matrix $P_i$ is composed of the camera's
intrinsic parameters, containing focal length and image center, as
well as extrinsic parameters which specify the rotation and
translation of the camera's position in 3D world coordinates at the
time that image $i$ is taken. These extrinsic parameters are
determined by the backpack hardware and localization algorithms
\cite{chen2010indoor, liu2010indoor, kua2012loopclosure} and are quite
noisy.

Because our backpack system takes photos at a rate of 5 Hz, thousands
of images are available for texturing each surface in our model. Our
goal in designing a texture mapping process is to decide which of
these images should be used, and where their contents should map onto
the final texture, in order to minimize any visual discontinuities or
seams that would suggest that the plane's texture is not composed of a
single continuous image.

@@

\subsection{Tile-based Texture Mapping}
\label{sec:tileBasedMapping}
Ignoring the fact that the camera matrices $P_{1..M}$ are inaccurate,
a simple texturing approach can be performed by discretizing the
target plane into small square tiles, in our case 5 pixels across, and
choosing an image to texture each tile.

@@

Once we have a list of candidate images for $t$, we define a scoring
function in order to objectively select the best image for texturing
$t$. Since resolution decreases and camera pose errors compound over
distance, we wish to minimize the distance between cameras and the
surfaces they texture. Additionally, we desire images that are
projected perpendicularly onto the plane, maximizing the resolution
and amount of useful texture available in their projections, as well
as minimizing any parallax effects due to real-world geometry not
accurately represented by our digital model. In other words, we wish
to minimize the angle between the tile's normal vector and the camera
axis for images selected for texture mapping. These two criteria can
be met by maximizing the function $\frac{1}{d} (-1 \cdot \vec{c})
\cdot \vec{n}$ as shown in Figure
\ref{fig:scoringFunction}. Specifically, $d$ is the distance between
the centers of a camera and a tile, and $\vec{n}$ and $\vec{c}$ are
the directions of the plane's normal and the camera axis respectively.

\begin{figure}
  \centering \subfloat[][]{\includegraphics[width=3in,
    height=1.2in]{wall2_naive.jpg}}
  \subfloat[][]{\includegraphics[width=3in,
    height=1.2in]{wall2_naive_shift.jpg}}

  \centering \subfloat[][]{\includegraphics[width=3in,
    height=1.2in]{wall2_cache_shift.jpg}}
  \subfloat[][]{\includegraphics[width=3in,
    height=1.2in]{wall2_shortest_shift.pdf}}

  \centering \subfloat[][]{\includegraphics[width=3in,
    height=1.2in]{wall2_cache_shift_blend.pdf}}
  \subfloat[][]{\includegraphics[width=3in,
    height=1.2in]{wall2_shortest_shift_blend.pdf}}
  \caption{(a) Tile-based texturing. (b) Tile-based texturing after
    image alignment. (c) Tile-based texturing after image alignment
    with caching. (d) Shortest path texturing after image
    alignment). (e) Same as (c) with blending. (f) Same as (d) with
    blending.}
  \label{fig:compareAll}
\end{figure}

As Figure \ref{fig:compareAll}(a) demonstrates, this approach leads to
the best texture for each tile independently, but results in many
image boundaries with abrupt discontinuities between tiles, due to
significant misalignment between images. Given our high number of input images, a better image selection procedure could be used, as described in Section \ref{sec:imageCompositing}, but a more significant and reliable improvement can be found through the alignment of our images.

@@

Feature detection and matching works best when multiple unique visual
references exist in the environment and are present within multiple
images. In contrast, our indoor environments have a high prevalence of
bare surfaces, as well as repeating textures that cause difficulty in
disambiguating features. This lack of reliable reference points often
results in errors when matching images together.

Additionally, our datasets often contain long chains of images, which
leads to error accumulation when image correspondences are not
accurate. For example, when matching a long chain of images through
homographies, a pixel in the $nth$ image is translated into the
first image's coordinates by multiplying by the $3\times3$ matrices $H_1
H_2 H_3 ... H_n$. Errors in any of these homography matrices are
propagated to all further images, resulting in drift. Figure
\ref{fig:mosaic3D}(a) shows the output of the AutoStitch software
package, which performs homography-based image mosaicing
\cite{autostitch}. Though AutoStitch performs well in areas with dense
features, it has difficulty with even short segments of bare walls,
and produces errors that cause straight lines to appear
distorted. Many areas with fewer visual features or repeating texture patterns failed
outright using AutoStitch.

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=6in]{autostitchResult.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=6in]{graphApproach.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=6in]{finalLong.jpg}}

  \caption{Texture alignment via (a) image mosaicing, (b) the
    graph-based localization refinement algorithm from
    \cite{chen2010indoor}, and (c) the method proposed in this paper.}
  \label{fig:mosaic3D}
\end{figure}

Image projections can also be aligned by iteratively adjusting camera
poses to maximize matches. This is generally done over 6 degrees of
freedom \cite{liu2010indoor}. The approach in the cited paper works
well locally, but over larger areas results again in error
accumulation and drift, as shown in Figure
\ref{fig:mosaic3D}(b). Furthermore, it adjusts image locations using
a long iterative process, which results in high runtime.

@@

All of our following calculations and alignments are performed in 2D,
partly for efficiency, partly because we have a large number of images
to choose from, and partly because the nature of our input data is
such that localization error chiefly occurs in two dimensions, in the
same plane as the surface being projected onto. Recall that our data
acquisition system is backpack-mounted, with cameras facing to the
sides of the operator. The operator makes efforts to walk parallel to
walls, and the localization and model-generation algorithms that
provide our input conform the operator's path to be straight and
parallel to detected surfaces \cite{kua2012loopclosure,
  sanchez2012point}. Since the operator walks in as upright a position
as possible, errors in roll, yaw, and the operator's distance from
parallel surfaces are minimal. Thus, our highest errors stem from
uncertainty in the operator's pitch, equivalent to rotation around the
camera axes, as well as location along surfaces. These equate to 2D
rotation and translation in a surface's plane.

\subsection{Geometry-based Alignment}
\label{sec:geometryAlignment}
After computing each image's projection onto the target surface, as
described in Section \ref{sec:simpleTextureMapping}, we detect line
segments in the image projections using Hough transforms. Experience
and intuition show that walls in indoor environments often contain
linear features that are either horizontal or vertical, corresponding
to doors, windows, posters, etc. Thus, when texturing walls, we rotate
images in 2D such that dominant lines are made to be horizontal or
vertical. This can be further extrapolated by orienting lines with a
wall's boundaries, for example in areas with a slanted roof.

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[height=1.5in]{geometryAlign_planes.png}}
  \subfloat[][]{\includegraphics[height=1.5in]{geometryAlignment.jpg}}
  \caption{(a) When texturing the red surface, it makes sense to align
    images to surface boundaries and intersections with other
    surfaces. (b) Red lines are geometry-based lines, while green
    lines are detected in the image via Hough transform. Above is the
    original projection using the input camera poses, and below is the
    projection after being adjusted for alignment.}
  \label{fig:geometryAlignment}
\end{figure}

At this point in time, image occlusions have not been accounted
for. As a result, some image projections contain texture that should
project to an adjacent surface, generally with a linear boundary where
the two surfaces meet. If this linear boundary is detected by Hough
transform, the image is rotated and shifted in 2D such that the visual
boundary between two surfaces in an image projection matches the
physical boundary in our digital model. An example of such an
adjustment is in Figure \ref{fig:geometryAlignment}

To perform the rotation and alignment, we use the RANSAC
\cite{fischler1981random} framework. RANSAC allows us to determine an
optimal rotation and translation to match up two groups of line
segments, while accounting for outliers. First, a rotation angle is
computed by gathering all pairs of image line segments and geometry
line segments with less than 20 degrees between them. We then run the
RANSAC algorithm by sampling 4 of these angles at a time, and using
their average as the fitting function. Angles over 2 degrees from this
average are considered outliers. Once this 2d rotation angle is
determined, it is applied to the image, and projections are
re-calculated and image-based lines are re-detected. Now, all pairs of
image line segments and geometry line segments with less than 0.2
degree difference and less than 250 mm distance at their furthest
points are gathered. RANSAC is then similarly employed to calculate
optimal horizontal and vertical translations independently, by
sampling 4 pairs at a time and considering an outlier threshold of 10
mm. If a translation is applied, we indicate that the image was
aligned to geometry in one or both dimensions, which will reduce
further shifting in Section \ref{sec:robustSIFTFeatureMatching}.

@@

\subsection{2D Feature Alignment}
\label{sec:robustSIFTFeatureMatching}
Our next step is to align overlapping images by searching for
corresponding points between all pairs of overlapping images. We use
SIFT features for their high detection rate, and choose to use feature
alignment rather than pixel or intensity-based alignment due to the
differences in lighting as well as possible occlusion among our
images, both of which feature alignment is less sensitive to
\cite{lowe1999object, mikolajczyk2005performance, szeliski2006image}.
We use SiftGPU \cite{siftgpu} for its high performance on both feature
detection as well as pairwise matching. These matches determine $d^x$
and $d^y$ distances between each pair of features for two image
projections, though these distances may not always be the same for
different features. Since indoor environments often contain repetitive
features such as floor tiles or doors, we need to ensure that
SIFT-based distances are reliable. First, we only perform alignment on
parts of images that overlap given the original noisy poses. Second,
we discard feature matches that correspond to an image distance
greater than 40 pixels from what the noisy poses estimate. In order to
utilize the remaining feature matches robustly, RANSAC
\cite{fischler1981random} is again used to estimate the optimal $d^x_{i,j}$
and $d^y_{i,j}$ distances between two images $i$ and $j$. For this
application, the RANSAC fitting function finds the average distance
between selected matches in a pair of images, and the distance
function for a pair of points is chosen to be the difference between
those points' SIFT match distance and the average distance computed by
the fitting function. We use a 10 pixel threshold, so that SIFT
matches are labeled as outliers if their horizontal or vertical
distances are not within 10 pixels of the average distance computed by
the fitting function.

We now use the RANSAC-calculated $d^x_{i,j}$ and $d^y_{i,j}$ distances
between each pair of images to refine their positions using weighted
linear least squares. There are a total of $M^{2}$ possible pairs of
images, though we only generate distances between images with SIFT
matches. Given these distances and the original image location
estimates, we can solve a least squares problem
($\textrm{min}_{\vec{\beta}} ||A \vec{\beta} - \vec{\gamma}||_2^2 $)
to estimate the location of the images on the plane. The
$M$-dimensional vector $\vec{\beta}$ represents the unknown $x$
location of each image on the plane for $1 \dots M$. The optimal $x$
and $y$ locations are obtained in the same way, so we only consider
the $x$ locations here:

@@

After completing this same process for the $y$ dimension as well, and
making the resultant shifts, our images overlap and match each other
with far greater accuracy. Applying the simple mapping scheme in
Section \ref{sec:tileBasedMapping} to the same wall used in that
section results in Figure \ref{fig:compareAll}(b), which has far fewer
discontinuities, though errors due to lighting differences and
repeating features are still visible.

@@

The result of this caching approach is shown in Figure
\ref{fig:compareAll}(c), where seams are now reduced as
compared to Figure \ref{fig:compareAll}(b).

@@

Figure \ref{fig:dagCreation} demonstrates the construction of a DAG
from overlapping images of a hallway wall. Images are sorted by
horizontal location left to right, and become nodes in a
graph. Directed edges are placed in the graph from left to right
between overlapping images. The weights of these edges are determined
by the SSD cost function. Next, we add two artificial nodes, one start
node representing the left border of the plane, and one end node
representing the right border of the plane. The left(right) artificial
node has directed edges with equal and arbitrary cost $C_0$ to(from)
all images that meet the left(right) border of the plane. We now solve
the shortest path problem from the start node to the end node. This
results in a set of images completely covering the plane horizontally,
while minimizing the cost of seams between images.

@@

After performing linear alpha blending across overlapping regions, our
texture mapping process is complete. Figures \ref{fig:compareAll}(e)
and \ref{fig:compareAll}(f) show the blended versions of Figures
\ref{fig:compareAll}(c) and \ref{fig:compareAll}(d) respectively. It
is clear that the shortest path approach exhibits much better
alignment and fewer seams than the tile-caching method, as Figure
\ref{fig:compareAll}(f) clearly has the best visual quality among the
textures in Figure \ref{fig:compareAll}.

\section{Results and Conclusions}
\label{sec:resultsAndConclusions}
Examples of ceilings and floors textured with the tile caching
approach, and walls textured with the shortest path approach, are
displayed in Figure \ref{fig:results}. High resolution texture
comparisons, as well as further examples and interactive walkthroughs are
available at \footnote{\url{http://www.eecs.berkeley.edu/~pcheng/textureMapping}}.

As mentioned earlier, our approach is quite efficient. The top wall in
Figure \ref{fig:results}(a) was generated with 7543 $\times$ 776
pixels, and spans a 40-meter long wall. Given 41000 input images, a
2.8GHz consumer-grade laptop takes approximately a minute to pick 36
candidate images, followed by another minute to perform both image
alignment and the shortest path texturing method, with the majority of
computation time spent calculating SIFT matches within the SiftGPU
framework. While not quite real-time visualization, the process is
capable of generating quick updates after changes in various
parameters or modifications to input data.

In this paper, we have developed an approach to texture mapping models
with noisy camera localization data. We are able to refine image
locations based on geometry references and feature matching, and
robustly handle outliers. Using the tile-based mapping approach, we
can texture both simple rectangular walls as well as complex floor and
ceiling geometry. We also implemented a shortest path texturing method
that produces seamless textures on planes where multiple head-on
images are available. Both of these approaches are highly modular, and
easily tunable for different environments and acquisition systems.

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=3in]{finalfloors.jpg}} ~~~~~~~~
  \centering
  \subfloat[][]{\includegraphics[width=3in]{finalceilings.jpg}}

  \centering \subfloat[][]{
    \includegraphics[height=2in, width=3in]{floorcropped.jpg} ~~~~~~~~
    \includegraphics[width=3in]{pier15floor.jpg}
  }

  \centering{
    \includegraphics[width=1.9in]{fullmodel.png}
    \includegraphics[width=1.9in]{threestoryfull2.png}
    \includegraphics[width=1.9in]{pier15.png} \\
    \includegraphics[width=6in]{threestoryfull.png}
  }

  \centering \subfloat[][]{ }
  \caption{Examples of our final texture mapping output for (a) walls,
    (b) ceilings, (c) floors, (d) full models.}
  \label{fig:results}
\end{figure}
